---
author: "Eric Hsienchen Chu"
output:
  html_document: default
  pdf_document: default
editor_options:
  markdown:
    wrap: sentence
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      error = TRUE, fig.align="center", fig.height = 5, fig.width = 8)
library(tidyverse)
library(stringr)
library(lubridate)

#library(plm)       # Panel data analysis library
#library(car)       # Companion to applied regression 
#library(tseries)   # For timeseries analysis
#library(lmtest)    # For hetoroskedasticity analysis
#library(directlabels)
#library(ggrepel)
#library(ggtext)
#library(gganimate)
#library(gifski)
#library(kableExtra)
#library(ggpubr)
#library(rddtools)
#library(magrittr)
#library(maps)
#library(mapdata)
#library(broom)  # Convert models to data frames
#library(rdrobust)  # For robust nonparametric regression discontinuity
#library(estimatr)  # Run 2SLS models in one step with iv_robust()
#library(ivreg)
#library(ivmodel)
#library(gtools)

# network
library(sna)
library(igraph)
library(threejs)
library(ggplot2)
library(ggraph)
library(ggrepel)
library(intergraph)
library(tidygraph)

#source("../viridis.R")
```

## [SNA LINKS 2023] 

#### Author: Eric Hsienchen Chu

#### Date: June 7, 2022


- - -

#### TA: Francisco Trincado-Munoz

#### SCRIPT Day3 - [Intermediate SNA with R (LINKS Workshop 2023](https://www.linksworkshop.org/intermediate-sna-r))

- - -

#### Case 1: Dataset is an edgelist 

```{r}
#Step 1 - Prepare your data

#The first dataset we “Dataset1-Media-Example-NODES.csv” and “Dataset1-Media-Example-EDGES.csv”

nodes = read.csv("Dataset1-Media-Example-NODES.csv", header = T, as.is = T)
links = read.csv("Dataset1-Media-Example-EDGES.csv", header = T, as.is = T)
```


```{r}
# To examine the data 
# use head or tail function
head(nodes)
head(links)

tail(nodes)
tail(links)
# If you have a bigger sample, you can specify the number of lines you want to see, default is 5
head(nodes, n = 7)
```

To convert the data into igraph object us the graph_from_data_frame() function. This function requires two sets of data:

(1) d: the edges of the network, it recognizes the two first columns as the IDs (source & target). 

  - The following columns are considered edge attributes (e.g., weight, type, etc.)
  
(2) vertices: The first column needs to be the nodes ID. All the following columns are nodes attributes. 

  - Nodes ID need to match with the IDs in the edgelist
  
(3) directed: you needs to specify "T" o "1" for directed or "F" or "0" for undirected


```{r}
# Step 2 - Get your data into igraph format 
#library(igraph)

net = graph_from_data_frame(d = links, vertices = nodes, directed = T)
net
```


The igraph object includes the following descriptions

> e.g. IGRAPH 3cbdde0 DNW- 17 49 --

(1) D or U, for directed or undirected graph

(2) N for named graph (where nodes have a name)

(3) W for a weighted graph (where edges have a "weight" attribute)

(4) B for a bipartite or two-mode network (where nodes have a "type" attribute)

The numbers represent the number of nodes & edges in the graph

The extra columns in the edgelist or NodesID files are considered attributes

- (g/c): graph-level character attribute

- (v/c): vertex level character attribute

- (e/n): edge-level numeric attribute


```{r}
# To access the nodes, edges or attributes you have to use the following codes
E(net)        #gives you the EDGES of the network
V(net)        #gives you the VTICES of the network
E(net)$type   #gives you the edge attribute "type"
V(net)$media  #gives you the node attribute "media"

E(net)$newVariable = ifelse(E(net)$type == "mention", 1, 0)

#Using this commands you can filter or request specific nodes or edges
#e.g.
V(net)[media == "BBC"]
E(net)[type == "mention"]


# To do the opposite procedure and get an edgelist from the igraph object
# use as_edgelist() function
as_edgelist(net, names = T)
matrix = as_adjacency_matrix(net, attr = "weight")

#class(as.array(matrix))
#write.csv(as.array(matrix), "matrix.csv")


as_data_frame(net, what = "edges")
as_data_frame(net, what = "vertices")

#################
```

- - -

#### Case 2: Dataset is an matrix

```{r}
##### Case 2.1 - You have the data ready

#Step 1 - Prepare your data

#The second dataset are “Dataset2-Media-Example-NODES.csv” and “Dataset2-Media-Example-EDGES.csv”
nodes2 = read.csv("Dataset2-Media-User-Example-NODES.csv", header = T, as.is = T)
links2 = read.csv("Dataset2-Media-User-Example-EDGES.csv", header = T, row.names = 1)

#Examine the data
head(nodes2)
head(links2)

#This matrix is a two-mode network
typeof(links2)

links2 = as.matrix(links2)

typeof(links2)

dim(links2)
dim(nodes2)
```


To read the matrix as an igraph object use the graph_from_incidence_matrix() function. To get a two mode network to work in igraph you need to set an attribute called "type" where type is "FALSE" or "0" for vertices in one mode and "TRUE" or "1" for vertices in the other mode.


```{r}
net2 = graph_from_incidence_matrix(links2)
net2
table(V(net2)$type)
```



```{r}
##### Case 2.2 - You have to put your data together

links3 = as.matrix(read.csv("BBT1_NN.csv", header = F))
links3

#write.csv(links2, "test.csv")

#To create a file with attributes

nodes3 = as.data.frame(LETTERS[1:nrow(links3)]) 

#nrow() gives you the number of rows
#ncol() gives you the number of columns
#alternatively you can use dim(links3)[1] for the rows 
# or dim(links3)[2] for the columns

# To change a column name use the names() function
names(nodes3) = "names"

#you can specify the specific column indicating the order
names(nodes3)[1] = "names"

#We can replaces the names for others
nodes3$names = c("Sheldon","Penny","Leonard","Raj","Howard","Bernadette","Amy","Stuart")

colnames(links3) = nodes3$names
rownames(links3) = nodes3$names
links3

net3 = graph_from_adjacency_matrix(links3, mode = "directed", weighted = TRUE)
net3
plot(net3)

# What if now I have attributes
atts3 = read.csv("BBT1_ATTs.csv", header = TRUE)

# Four new variables
# (1) PhD : if the character has a PhD
# (2) Physics : if the character is in the Physics field
# (3) Gender :  MALE or FEMALE
# (4) Episodes : Number of Episodes the character appears in the series

# Remember if you want to add these variables to an existing data.frame you can use cbind() function
nodes3 = cbind(nodes3, atts3)

#Let's check 
nodes3$Gender
#How many Males and Females
table(nodes3$Gender)

#Average of episodes
mean(nodes3$Episodes,na.rm = T) #na.rm=T in case that you have NAs
round(mean(nodes3$Episodes,na.rm = T), 0)

#Let's put these variables in a igraph object
V(net3)$Gender = nodes3$Gender
V(net3)$PhD = nodes3$PhD
V(net3)$Physics = nodes3$Physics
V(net3)$Episodes = nodes3$Episodes

E(net3)$var = 1

#Check net3 again
net3
```


```{r}
##############################################
#Alternatively you set everything from the beginning

links3_Edgelist = as_edgelist(net3,names = T)
 
net3.1 = graph_from_data_frame(links3_Edgelist, directed = TRUE, vertices = nodes3)
net3.1
plot(net3.1)

##############################################
```



```{r}
# what if I want to create variables or assign values, some examples:

#(1) New variable for PhD in Physics
V(net3)$PhD_in_Physics =V(net3)$PhD*V(net3)$Physics

#(2) We want to know who was part of Season 3  #S1:17 eps #S2:23 eps #S3:23 eps
# ifelse() function ifelse(condition, TRUE,FALSE)
V(net3)$Season3 = ifelse(V(net3)$Episodes > 63, 1,0)

# color
color = c("orange","skyblue")

plot(net3,
     vertex.color = color[V(net3)$PhD_in_Physics+1], edge.arrow.size = .4)
     

V(net3)$size = ((V(net3)$Episodes/100)^3)+5

plot(net3,
     vertex.color = color[V(net3)$PhD_in_Physics+1], edge.arrow.size = .4)
```


#### What if I want to subset an ego network or data with $x$ characteristics


```{r}
#(1) induced_subgraph() function

net3.2 = induced_subgraph(net3,V(net3)$Season3 == 1, impl = "copy_and_delete")
net3.2

net3.2.1 = induced_subgraph(net3,V(net3)$name %in% c("Raj", "Sheldon", "Penny", "Howard"), impl = "copy_and_delete")

plot(net3.2, edge.arrow.size = .4)
plot(net3.2.1, edge.arrow.size = .4) 

#(2) delete.edges or delete.vertices
net3.3 = delete.vertices(net3,V(net3)$Gender == "Female")
net3.3
plot(net3.3, edge.arrow.size = 0.4)


#(3) neighbours
Ego_Sheldon = igraph::neighbors(net3, V(net3)["Sheldon"], mode="all")
Ego_Sheldon
plot(Ego_Sheldon)

Ego_Sheldon = V(net3)["Sheldon", .nei("Sheldon")] %>% 
  induced_subgraph(graph = net3) 
Ego_Sheldon

plot(Ego_Sheldon)
```

#### Changing from igraph to sna


```{r}
# igraph to network object

#install.packages('Intergraph')
library(intergraph)

# Check the network object
net3
class(net3)

#Two main functions asNetwork & asIgraph


net3_Network = asNetwork(net3)
class(net3_Network)

net3_Network

# Density
sna::gden(net3_Network)
# Reciprocity

sna::grecip(net3_Network,
            measure = c("edgewise"))

# Why the value is 1?

# Transitivity
sna::gtrans(net3_Network)

# =y to be used with the ergm package
#check more here: https://cran.r-project.org/web/packages/intergraph/vignettes/howto.html

# The other way around

net3_Igraph = asIgraph(net3_Network)
class(net3_Igraph)

net3_Igraph

V(net3_Igraph)$name = V(net3_Igraph)$vertex.names
net3_Igraph
```


#### Other useful tools for data management 

```{r}
# We will work with the data of Game of Thrones Season 3 (collected by Andre Beveridge of Macalester University). This is based on Season 3.
# 123 characters and 500 edges based on co-ocurrences 

GoT_Nodes = read.delim("GOT_Nodes.csv",sep = ",")
GoT_Edges = read.csv("GOT_Edges.csv",header = T)

head(GoT_Nodes)
head(GoT_Edges)

library(dplyr)
library(stringr)

# dplyr package is a really good ally to prepare data
# dplyr only works with data frames so you can use as.matrix() & as.data.frame() to convert them

# (1) Text operations 
# Using the stringr package you can use some text cleaning, always necessary
# the function mutate from dplyr changes values of columns or add new columns

GoT_Nodes_1 = GoT_Nodes %>%
    mutate_if(is.character,str_trim) %>% #to delete extra spaces before and after #str_squish() delete extra spaces between words
    mutate(name = str_to_title(name)) %>% # to set as capital only the first letter
    mutate(name = str_replace_all(name," ","_")) %>% #str_replace_all() replaces all the characters in the condition
    mutate(house = str_to_title(house))  #str_to_upper() & #str_to_lower

#(2) Selecting variables ]

GoT_Nodes_1 = GoT_Nodes_1 %>%
    select(name,house) #select() changes whether a column is selected or not

#(3) Filter rown based on values
GoT_Nodes_2 = GoT_Nodes_1 %>%
    filter((house != "Other"))

GoT_Nodes_2
```


```{r}
# When numeric, you can use == for equal or != for different
# or you can use %in% when is a character
# to filter many characters you can use  %in% c("Filter1","Filter2",etc...)

#Other useful dplyr functions
#rename() changes the names of columns
#relocate() changes the order of the columns
#summarise() collapses date to a group value

GoT_House = GoT_Nodes_2 %>%
  group_by(house) %>%
  summarise(n = length(house)) 

GoT_House

# Now before we carry on, let's do the same for the edges 
names(GoT_Edges)[1]="source" #Change the name of the first column

GoT_Edges_1 = GoT_Edges %>%
  mutate_if(is.character,str_trim) %>% #to delete extra spaces before and after #str_squish() delete extra spaces between words
  mutate(source = str_to_title(source)) %>% 
  mutate(source = str_replace_all(source," ","_")) %>%
  mutate(target = str_to_title(target)) %>% 
  mutate(target = str_replace_all(target," ","_"))
  
#(4) what if I only want the characters that have a house

GoT_Edges_2 = GoT_Edges_1 %>%
  filter(source %in% unique(GoT_Nodes_2$name)) %>%
  filter(target %in% unique(GoT_Nodes_2$name))


#Ready! Let's plot both networks with all the characters and only with those in a house

# A. All the characters

GoT_n_1=graph_from_data_frame(GoT_Edges_1, directed = TRUE, vertices = GoT_Nodes_1)
GoT_n_1
plot(GoT_n_1, edge.arrow.size = .4)


# B. Only characters with a house

GoT_n_2=graph_from_data_frame(GoT_Edges_2, directed = TRUE, vertices = GoT_Nodes_2)
GoT_n_2
plot(GoT_n_2,edge.arrow.size = .4)

#Let's color the houses
#install.packages('RColorBrewer')
library(RColorBrewer)

display.brewer.all()

display.brewer.pal(8, "Set3")

HouseColor =brewer.pal(nrow(GoT_House), "Set3")

#To match color and house we will do a trick
#Create a variable in the group dataset with the order for each house
GoT_House$Order =1:nrow(GoT_House)

#Now I will create a loop
for(i in 1:47)
{
  V(GoT_n_2)$Color[i] = HouseColor[GoT_House[GoT_House$house == V(GoT_n_2)$house[i],"Order"][[1]]]
}

V(GoT_n_2)$Color

plot(GoT_n_2,edge.arrow.size = .4, vertex.color = V(GoT_n_2)$Color,vertex.size = 10,vertex.label.cex = 0.8)

legend(x=-1.5, y=-0.5, GoT_House$house, pch=21,
       col="#777777", pt.bg=HouseColor, pt.cex=2, cex=.8, bty="n", ncol=1)
```

- - - 

#### Visualization

```{r}
# ggplot is an alternative with better visualization tools
library(ggplot2)
library(ggraph)
library(ggrepel)

ggraph(GoT_n_2) + 
  geom_edge_link() + 
  geom_node_point(aes(color = factor(house)))

# Because we do not specify layout, it uses the default: "auto"

# we will pick a specific layout: Kamada-Kawai layout algorithm
ggraph(GoT_n_2, layout = 'kk') +   #maxiter = 200
  geom_edge_link() + 
  geom_node_point(aes(color = factor(house)))

# You can also create a layout
#layout = create_layout(graph, layout = 'drl')

E(GoT_n_2)$Close = ifelse(E(GoT_n_2)$Weight > mean(E(GoT_n_2)$Weight,na.rm = T),"Close","Distance")
GoT_n_2

# If we want to create a circular plot

ggraph(GoT_n_2, layout = 'linear') + 
  geom_edge_arc(aes(colour = factor(Close)))+
  geom_node_point(aes(color = factor(house)))

ggraph(GoT_n_2, layout = 'linear', circular = T) + 
  geom_edge_arc(aes(colour = factor(Close)))+
  geom_node_point(aes(color = factor(house)))

# Change layout to Fruchterman-Reingold layout algorithm

ggraph(GoT_n_2,layout = "fr") + 
  geom_edge_link(aes(colour = factor(Close))) + 
  geom_node_point(aes(color = factor(house))) +
  theme_graph()

# Add labels to nodes
ggraph(GoT_n_2,layout = "fr") + 
  geom_edge_link(aes(colour = factor(Close))) + 
  geom_node_point(aes(color = factor(house))) +
  theme_graph()+
  geom_node_text(aes(label = name), repel=TRUE)

# Add a different shape
ggraph(GoT_n_2,layout = "fr") + 
  geom_edge_link(aes(colour = factor(Close))) + 
  geom_node_point(aes(color = factor(house), shape = factor(house))) +
  theme_graph()+
  geom_node_text(aes(label = name), repel=TRUE)

# If I want the size and transparency of the node to depend on the degree
ggraph(GoT_n_2,layout = "fr") + 
  geom_edge_link(aes(colour = factor(Close))) + 
  geom_node_point(aes(color = factor(house),size = degree(GoT_n_2), alpha = degree(GoT_n_2))) +
  theme_graph()+
  geom_node_text(aes(label = name), repel=TRUE)

# You can also facet the plots depending on the ties
ggraph(GoT_n_2,layout = "fr") + 
  geom_edge_link(aes(colour = factor(Close))) + 
  geom_node_point(aes(color = factor(house))) +
  theme_graph()+
  geom_node_text(aes(label = name), repel=TRUE) +
  facet_edges(~Close) +
  labs(color="House", edge_colour = "Relationship")
  

# Let's add more attributes

Got_Plot1 = ggraph(GoT_n_2, "fr")+
  geom_edge_link(aes(edge_width = Weight),alpha = 0.7)+ # filter = (house == "Stark")
  geom_node_point(aes(color = factor(house),size = degree(GoT_n_2)), shape = 16)+
  geom_node_text(aes(label = name, size = 5),repel=TRUE)+
  scale_edge_width(range = c(0.1,3))+
  labs(color="House", size="Appearance", edge_width = "Co-appearance")+
  theme_graph() #legend.position = "none" if you do not want legend

#warnings()


Got_Plot1


ggsave(Got_Plot1, filename = "Got_Plot1.pdf", device = cairo_pdf,
       width = 30, height = 25, units = "cm", dpi=300)

ggsave(Got_Plot1, filename = "Got_Plot1.png",
       width = 30, height = 25, units = "cm", dpi=300)


# ggraph/ggplot2 can be integrated to dplyr


GoT_House = GoT_Nodes_2 %>%
  group_by(house) %>%
  summarise(n = length(house)) 

GoT_House
```


```{r}
#install.packages('tidygraph')
library(tidygraph)

# First you need to create a table, use as_tbl_graph():

GoT_n_2_Table = as_tbl_graph(GoT_n_2)
GoT_n_2_Table

# To change nodes or attributes we have to use activate(nodes) or activate(edges)
GoT_n_2_Table %>%
  activate(nodes) %>%
  mutate(degree = centrality_degree())


# We can use other ggplot graphs
GoT_n_2_Table %>%
  activate(nodes) %>%
  mutate(cl = centrality_closeness()) %>%
  as_tibble() %>%
  ggplot() +
  geom_histogram(aes(x=cl), bins = 30) +
  theme_bw()

#?centrality_closeness
```



```{r}
# We can check the data as table:
GoT_n_2_Table %>%
  activate(nodes) %>%
  mutate(cl = centrality_closeness()) %>%
  as_tibble() %>%
  select(name,cl) %>%
  arrange(desc(cl))


GoT_n_2_Table %>%
  activate(edges) %>% # You can also activate nodes
  filter(Close == "Close") %>%
  ggraph(layout = "fr") + 
  geom_edge_link() + 
  geom_node_point(aes(color = factor(house))) +
  geom_node_text(aes(label = name), repel=TRUE) +
  theme_graph()
```



```{r}
# To read spss files

#library(foreign)

#SPSS_Data= read.spss(file ="FileName.sav", to.data.frame= TRUE, reencode = TRUE)
```



```{r}
data(GoT_n_2)
graphjs(GoT_n_2, bg="black", vertex.size = 0.5)

data(GoT_n_1)
graphjs(GoT_n_1, bg="black", vertex.size = 0.5)

```


