---
author: "Eric Hsienchen Chu"
output:
  html_document: default
  pdf_document: default
editor_options:
  markdown:
    wrap: sentence
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      error = TRUE, fig.align="center", fig.height = 6, fig.width = 8)
library(tidyverse)
library(stringr)
library(lubridate)

#library(plm)       # Panel data analysis library
#library(car)       # Companion to applied regression 
#library(tseries)   # For timeseries analysis
#library(lmtest)    # For hetoroskedasticity analysis
#library(directlabels)
#library(ggrepel)
#library(ggtext)
#library(gganimate)
#library(gifski)
#library(kableExtra)
#library(ggpubr)
#library(rddtools)
#library(magrittr)
#library(maps)
#library(mapdata)
#library(broom)  # Convert models to data frames
#library(rdrobust)  # For robust nonparametric regression discontinuity
#library(estimatr)  # Run 2SLS models in one step with iv_robust()
#library(ivreg)
#library(ivmodel)
#library(gtools)

# network
library(sna)
library(igraph)
library(threejs)
library(ggplot2)
library(ggraph)
library(ggrepel)
library(intergraph)
library(tidygraph)
library(xUCINET)

#source("../viridis.R")
```

## [SNA LINKS 2023] 

#### Author: Eric Hsienchen Chu

#### Date: June 7, 2022


- - -

#### TA: Francisco Trincado-Munoz

#### SCRIPT Day4-1 Two-Mode - [Intermediate SNA with R (LINKS Workshop 2023](https://www.linksworkshop.org/intermediate-sna-r))

- - -

#### I. Preparing the data

```{r}
# From the class
# Network of 6 scientists and 8 publications

MATRIX68 = matrix(c(1,0,0,0,0,0,0,0,
                     +                    1,0,0,1,0,0,0,0,
                     +                    1,1,0,0,0,0,0,0,
                     +                    1,1,1,0,0,0,0,0,
                     +                    0,0,1,1,1,1,1,1,
                     +                    0,0,0,0,0,0,1,1),
                     +                    6,8,byrow = TRUE)
rownames(MATRIX68) = c("a1","a2","a3","a4","a5","a6")
colnames(MATRIX68) = c("e1","e2","e3","e4","e5","e6","e7","e8")
```


```{r}
# Plotting
gplot(MATRIX68, displaylabels=TRUE, usearrows=FALSE, gmode="twomode")

# Row sums (degree of actors?)
rowSums(MATRIX68)

# Col Sums (degree of events?)
colSums(MATRIX68)

# Proportion for actors
rowSums(MATRIX68)/NCOL(MATRIX68)

# Proportion for events
colSums(MATRIX68)/NROW(MATRIX68)
```


```{r}
# Total Contribution 
t(t(MATRIX68)/colSums(MATRIX68))

rowSums(t(t(MATRIX68)/colSums(MATRIX68)))

# Average Contribution
rowSums(t(t(MATRIX68)/colSums(MATRIX68)))/(rowSums(MATRIX68))
```

- - - 

#### II. One-mode Projections


```{r}
# Projections
(Actors_Projection = MATRIX68 %*% t(MATRIX68)) # A*(A^T)
# You can read this as papers that actors collaborate
# The diagonal is the number of papers by author
diag(Actors_Projection)
```


```{r}
(Events_Projection = t(MATRIX68) %*% MATRIX68)
# Number of common collaborators between papers

# The diagonal represents the total authors by paper
diag(Events_Projection)
```


> For the analysis, you usually have to ignore the diagonal and set this to zero

```{r}
diag(Actors_Projection) = 0
diag(Events_Projection) = 0

par(mar = c(0, 0, 0, 0))
gplot(Actors_Projection, displaylabels=TRUE,
      usearrows=FALSE, edge.lwd=1)

par(mar = c(0, 0, 0, 0))
gplot(Events_Projection, displaylabels=TRUE,
      usearrows=FALSE, edge.lwd=1)
```

- - - 

#### Two-mode Network


```{r} 
# require packages
#library(xUCINET)
#library(sna)
#library(igraph)
```


```{r}
# Two-mode Network (what are they?), why?

# Let's have a look at one of the most popular two-mode dataset
# Southern women attendance to events
Davis_SouthernWomen

# The ties represent attendance at 14 social events by 18 Southern women

#Check the matrix only
Davis_SouthernWomen$Attendance
```

#### STEP 1 - Plotting 

```{r}

# (because plotting always help to understand the network)

# Plotting with igraph

BN_SW_IG = graph_from_incidence_matrix(Davis_SouthernWomen$Attendance)

#Check the network

BN_SW_IG

#Check the type: FALSE (women) & TRUE (events)

table(V(BN_SW_IG)$type)

# We set some attributes for visualization
# Women are blue squares, events nodes are orange circles:

V(BN_SW_IG)$color = c("steel blue", "orange")[V(BN_SW_IG)$type + 1] #We add 1 cuz the type is 1 or 0 (TRUE or FALSE)  
V(BN_SW_IG)$shape = c("square", "circle")[V(BN_SW_IG)$type + 1]

#For the labels 
V(BN_SW_IG)$label <-c(substr(rownames(Davis_SouthernWomen$Attendance),1,2),colnames(Davis_SouthernWomen$Attendance))

# Igraph has a special layout for bi-partite networks, let's try it
```


```{r}
#w/o labels
plot(BN_SW_IG, vertex.label=NA, vertex.size=7, layout=layout_as_bipartite)

# w labels
par(mar = c(1,1,1,1))
plot(BN_SW_IG, vertex.size=7, layout=layout_as_bipartite, vertex.label.dist = 1.5,vertex.label.degree = -pi/2)
```


#### STEP 2 -  Analysing the network

In general, there are two options to analyze a bipartite network

- Option 1: Build a projection of the network and analyze each level separately

- Option 2: Analyze the whole network, without distinction

Let's check some things that we could do...


> `OPTION 1: Projections`

```{r}
# Using the xUCINET
# The xTwoModeToOneMode() function turns the data from two-mode to one-mode network

SW_PROJ_WOMEN <- xTwoModeToOneMode(Davis_SouthernWomen$Attendance)
SW_PROJ_WOMEN

# The result is a matrix with the co-attendance of the women to the events
# rows and columns contain the women names
# The value of the cell i,j represent the co-attendance to events of the woman i & the woman j
# The diagonal contains the total number of events that each woman attended

SW_PROJ_EVENTS <- xTwoModeToOneMode(t(Davis_SouthernWomen$Attendance))
SW_PROJ_EVENTS


# Plotting these networks
# We dichotomize at 2, higher than 2 events gets a tie

sna::gplot(SW_PROJ_WOMEN > 2,
           gmode="graph", 
           displaylabels = TRUE,         #Labels
           label.cex=.7,                 #Labels
           edge.lwd=SW_PROJ_WOMEN*1.5-4, #Edges
           vertex.cex=.7,                #Vertices
           vertex.col=7,                 #Vertices
           edge.col = "grey65")
```


```{r}
## Projection with normalization

# why do you need to normalize two-mode network?

# To use Bonacich’s normalization method for a one-mode projection, 
# we add the argument (Measure=“Bonacich”). We also multiply the value by 100.

SW_PROJ_EVENTS_NORM<-xTwoModeToOneMode(t(Davis_SouthernWomen$Attendance),Measure="Bonacich")*100

round(SW_PROJ_EVENTS_NORM)

#Plotting the resulting network

sna::gplot(SW_PROJ_EVENTS_NORM > 50, 
           gmode="graph", 
           displaylabels = TRUE,
           label.cex=.7,
           label.pos=5,
           label.col = 4,
           edge.lwd=(SW_PROJ_EVENTS_NORM>50)*(SW_PROJ_EVENTS_NORM-45)/14,
           vertex.sides=4,
           vertex.cex=1.9,
           vertex.rot=45,
           vertex.col=3)

#### You can use both projections to analyze centrality
```


```{r}
#############################################
# Alternatively

# Co-attendance of women
WOM_PROJ<- Davis_SouthernWomen$Attendance %*% t(Davis_SouthernWomen$Attendance) 
WOM_PROJ

# Checking the projection results
WOM_PROJ==xTwoModeToOneMode(Davis_SouthernWomen$Attendance)

# Co-attendance of events
EV_PROJ<-t(Davis_SouthernWomen$Attendance) %*% Davis_SouthernWomen$Attendance
EV_PROJ

##############################################
```


#### III. Two mode to Bipartite network


```{r}
xTwoModeToBipartite(Davis_SouthernWomen$Attendance)

# In this network women and events are in the columns and rows, the blocks in the diagonal contain zeros

DIM1<-dim(Davis_SouthernWomen$Attendance)
DIM1
# 18 14  # 18 women and 14 events

#To be able to use different colors and shapes, for the one mode and the second mode, we create a vector with 0 for actors and 1 for events. 

RC1<-c(rep(0, DIM1[1]),rep(1, DIM1[1]))

RC1 #the zero are for the women and ones are for the events

LABELS1<-c(substr(rownames(Davis_SouthernWomen$Attendance),1,2),colnames(Davis_SouthernWomen$Attendance))
LABELS1

#Now let's draw the network:
sna::gplot(Davis_SouthernWomen$Attendance, 
           mode = "fruchtermanreingold", 
           gmode="twomode", 
           #Labels
           displaylabels=TRUE, 
           label = LABELS1, 
           label.cex=.7,
           label.pos=5,
           #Edges
           usearrows=FALSE,
           #Vertices
           vertex.cex=RC1*.5+1.7, 
           vertex.rot=45, 
           vertex.col=RC1*7)
```

> `OPTION 2: Analyzing the whole network` 

```{r}
# Option 2.1: use the bipartite network and use classic one-mode analysis
# Option 2.2: some packages as xUCINET or tnet allow you to analyse two-mode and consider differences

####### (a) Degree Centrality

##### To calculate the degree centrality for the women

xDegreeCentrality(Davis_SouthernWomen$Attendance)

# *The max number of events is 14, the nDegree is the value/max(N_Events)

##### To calculate the degree centrality for the events, we simply transpose the matrix:

xDegreeCentrality(t(Davis_SouthernWomen$Attendance))


####### (b) Other centrality measures with xUCINET

# We can use the bipartite network and calculate closeness centrality
# However, we need to know that the normalization will be incorrect

BIP<-xTwoModeToBipartite(Davis_SouthernWomen$Attendance)

xClosenessCentrality(BIP)


####### (c) Global group measures with xUCINET

# xDensity() function can take a two-mode network as input

xDensity(Davis_SouthernWomen$Attendance)


BC<-xBiCliques(Davis_SouthernWomen$Attendance)
BC

BC%*%t(BC)   # is simple one-mode projection


xHierarchicalClustering(BC%*%t(BC), Input="Similarities", Method="average")
```


```{r}
############ Alternatively you can use igraph ##################

##### we will carry on with the same data for better understanding 

# we previously built an igraph object of the Davis Suthern women dataset
BN_SW_IG <- graph_from_incidence_matrix(Davis_SouthernWomen$Attendance)
BN_SW_IG
```

> `Option 2.1 - do it as normal`

```{r}
types <- V(BN_SW_IG)$type  ## getting each vertex "type" helps to order things
deg <- igraph::degree(BN_SW_IG)
bet <- igraph::betweenness(BN_SW_IG)
clos <- igraph::closeness(BN_SW_IG)
eig <- eigen_centrality(BN_SW_IG)$vector

cent_df_NOCHANGE <- data.frame(types, deg, bet, clos, eig)

cent_df_NOCHANGE<-cent_df_NOCHANGE[order(cent_df_NOCHANGE$type, decreasing = TRUE),] ## sort w/ `order` by `type`
#View(cent_df_NOCHANGE)
```

> `Option 2.2 - analyze each mode separately`

Tore Opsahl has a website about the analysis of bipartite networks and he developed the package tnet with some alternatives [https://toreopsahl.com/](https://toreopsahl.com/)

```{r}
install.packages("tnet", dependencies=TRUE)
library(tnet)  # start tnet

#tnet requires an edgelist object without names (only numbers) so we need to transform the igraph object

WoEv<-get.edgelist(BN_SW_IG, names=FALSE)

# In this case the women are in the 1st column and the events in the 2nd
head(WoEv)

# To get the events first we just alter the order of the columns
EvWo <- WoEv[, c(2, 1)]
head(EvWo) 

# Why this is important? Because the degree function is only calculated for the data in the 1st column

deg_WoEv <- degree_tm(WoEv)
deg_WoEv

deg_EvWo <- degree_tm(EvWo)
deg_EvWo

# Up to now, tnet only has the degree function to analyze bi-partite data
# However, Tore recommends to use the weighted projections to analyze the data

# We previously built the projections

EV_PROJ

# We set the diagonal to zero (to avoid self loops)
diag(EV_PROJ) <- 0

# Using as.tnet() function we get a numerical edgelist 
EV_PROJ_TN <- as.tnet(EV_PROJ)

# Warning message:
#   In as.tnet(SW_PROJ_EVENTS_NORM) :
#   Data assumed to be weighted one-mode tnet (if this is not correct, specify type)
# This is just a warning 

head(EV_PROJ_TN)
# sender(i) receiver(j) weight

Events_Wdeg <- degree_w(EV_PROJ_TN)
Events_Wbet <- betweenness_w(EV_PROJ_TN)
Events_Wclos <- closeness_w(EV_PROJ_TN, gconly=FALSE)

# Note: tnet does not include eigenvector centrality

Events_W_cent_df <- data.frame(Events_Wdeg, Events_Wbet, Events_Wclos)

Events_W_cent_df
```


```{r}
##### we could do the same for the women projection

# We previously built the projections

WOM_PROJ

# We set the diagonal to zero (to avoid self loops)
diag(WOM_PROJ) <- 0

# Using as.tnet() function we get a numerical edgelist 
WOM_PROJ_TN <- as.tnet(WOM_PROJ)

# Warning message:
#   In as.tnet(SW_PROJ_EVENTS_NORM) :
#   Data assumed to be weighted one-mode tnet (if this is not correct, specify type)
# This is just a warning 

head(WOM_PROJ_TN)
# sender(i) receiver(j) weight

Women_Wdeg <- degree_w(WOM_PROJ_TN)
Women_Wbet <- betweenness_w(WOM_PROJ_TN)
Women_Wclos <- closeness_w(WOM_PROJ_TN, gconly=FALSE)

# Note: tnet does not include eigenvector centrality

Women_W_cent_df <- data.frame(Women_Wdeg, Women_Wbet, Women_Wclos)
rownames(Women_W_cent_df)<-rownames(Davis_SouthernWomen$Attendance)
Women_W_cent_df
```






